\textbf{Computational Fuzzy Extractors}

### 2013-05-17 trunk/CompFuzzExt.tex
clean up 5
further edits to intro -- in light of the talk

\paragraph {Our Negative Results}
We first study whether it could be fruitful to relax the definition of the main building block of a fuzzy extractor, called \emph{secure sketch}.  A secure sketch provides information reconciliation: it  produces a public value $s$ that allows recovery of $w$ from any close value $w'$.  The traditional secrecy requirement of a secure sketch is that $w$ has high min-entropy conditioned on $s$.  This allows the fuzzy extractor of~\cite{DBLP:journals/siamcomp/DodisORS08}  to form the key $r$ by applying a randomness extractor~\cite{nisan1993randomness} to $w$, because randomness extractors get random strings from strings with conditional min-entropy.

@@

By considering this computational secrecy requirement, we construct the first \emph{lossless} (computational) fuzzy extractors, where the derived key $r$ is as long as the entropy of the source $w$. % without conditioning on anything---i.e., before $\gen$ is run (see~\thref{thm:lossless secure conductor log})).  
Our construction of lossless fuzzy extractors~(\consref{cons:informal construction}) is based on the Learning with Errors~(LWE) assumption due to Regev~\cite{regev2005LWE, regevLWEsurvey},  which says that decoding a random linear code is computationally difficult.  
%We show that this relaxation of the security requirement is  useful.  Specifically, we also define the notion of computational fuzzy extractors, in which the extracted key is required to be pseudorandom---i.e., to look random to any computationally bounded observer.
%We show that our relaxed notion of secure sketches is sufficient to construct computational  fuzzy extractors via a simple and efficient construction, utilizing reconstructive extractors~\cite{barak-computational}.  The resulting fuzzy extractor produces a key that is as long as the entropy of $w$, minus only the loss necessary in any extractor construction.
%\lnote{it would be good to find another application} 
Our lossless fuzzy extractor is built via the code-offset construction~\cite{JW99},\cite[Section 5]{DBLP:journals/siamcomp/DodisORS08} used in prior work.  To apply the LWE assumption, we correct errors of a random linear code, which limits our decoding algorithm to reconciling a logarithmic number of differences.

The recent result of D\"{o}ttling and M\"{u}ller-Quade~\cite{dottling2012} shows the hardness of LWE when the error vector comes from the uniform distribution, with each coordinate ranging over a small interval.  This allows us to directly use our source $w$ as the LWE error vector.  To utilize the result of~\cite{dottling2012}, we need to assume that the input comes from the uniform distribution. We are able to relax this limitation and allow $w$ to come from a symbol-fixing source~\cite{KZ07} (each dimension is either uniform or fixed). This relaxation requires new results about the hardness of LWE when samples have a fixed~(and adversarially known) error vector, which may be of independent interest~(\thref{thm:blockLWE}).

\paragraph{An Alternative Approach}  Computational extractors~\cite{krawczyk2010cryptographic, barak2011leftover, dachman2012computational} have the same goal of obtaining a pseudorandom key $r$ from a source $w$ in the setting without errors.  They can be constructed, for example, by applying to pseudorandom generator to the output of an information-theoretic extractor.  One way to build a computational \emph{fuzzy} extractor is by using a computational extractor instead  of the information-theoretic extractor in the construction of  \cite{DBLP:journals/siamcomp/DodisORS08}.   However, this approach is possible only if conditional min-entropy of $w$ conditioned on the sketch $s$ is high enough.  
%We ask whether it is possible to avoid this loss by making whole process ``computational.''
We compare the two approaches in \secref{sec:prg based comparison}.

@@

%In \secref{sec:defCompFuzzyExtractors} we will use a slightly weaker notion, in which \rec is allowed to fail with small probability.\lnote{Can we add a comment that this corresponds to errors that can depend on the input but not the sketch? It's not obvious from the way this is written.} \lnote{can we just have a single definition --- the one below --- instead of two?}
%\begin{definition}
%We say that pair of procedures $(\sketch, \rec)$ is $(\mathcal{M}, m, \tilde{m}, t, \epsilon)$ is an \emph{approximately correct secure sketch} if \defref{def:secure sketch} holds with correctness replaced with the following property, $\forall w, w'$ such that $\dis(w, w')\leq t$:
%\[
%\Pr[\rec(w', \sketch(w)) = w]\geq 1-\epsilon.
%\] where the probability is taken over the coins of \sketch and \rec.
%\end{definition}
\begin{lemma}%\protect{\cite[Lemma 4.1]{DBLP:journals/siamcomp/DodisORS08}}
\label{lem:fuzzy ext construction}
Assume $(\sketch, \rec)$ is an $(\mathcal{M}, m, \tilde{m}, t)$-secure sketch with error $\delta$, and let $\ext:\mathcal{M}\times \zo^d \rightarrow \zo^\ell$ be a $(\tilde{m}, \epsilon)$-average-case extractor.  Then the following $(\gen, \rep)$ is an $(\mathcal{M}, m, \ell, t, \epsilon)$-fuzzy extractor with error $\delta$:
\begin{itemize}
\item $\gen(w):$ set $P=(\sketch(w), x), R=\ext(w;x)$, and output $(R,P)$.
\item $\rep(w', (s, x)):$ recover $w=\rec(w',s)$ and output $R=\ext(w;x)$.
\end{itemize}
\end{lemma}
The main parameter we will be concerned with is the entropy loss of the construction.  In this paper, we ask whether a smaller entropy loss can be achieved by considering a fuzzy extractor with a computational security requirement.  We relax the security requirement of \defref{def:fuzzy extractor} to require a pseudorandom output instead of a truly random output.

@@

%=============================================================
\section{Impossibility of Computational Secure Sketches}
\label{sec:defCompFuzzyExtractors}
In this section, we consider whether it is possible in build a secure sketch that retains significantly more computational than information-theoretic entropy.  We consider two different notions for computational entropy, and for both of them show that corresponding secure sketches are subject to the same upper bounds as those for information-theoretic secure sketches. Thus, it seems that relaxing security of sketches from information-theoretic to computational does not help.

@@

\begin{theorem}
\label{thm:imp of unp entropy}
Let $Z$ be an alphabet $(\sketch, \rec)$ be an unpredictability-entropy $(Z^n, m, \tilde{m}, t)$-secure sketch that is $(\epsilon, s_{sec})$-secure with error $\delta$.  If $s_{sec} \geq t(|\rec|+n\log |Z|)$, then $\tilde{m}\leq n\log |Z| - \log |B_t(\cdot)| + \log(1-\epsilon -t\delta)$.
\end{theorem}
In particular, if the input is uniform, the entropy loss is about $\log |B_t(\cdot)|$.  As mentioned at the beginning of~\secref{sec:defCompFuzzyExtractors}, essentially the same entropy loss can be achieved with information-theoretic secure sketches, by using the randomized code-offset construction. However, it is conceivable that unpredictability entropy secure sketches could achieve lower entropy loss with greater efficiency for some parameter settings.

@@

In this section we describe our main construction.  In this section, we consider only a uniform source $W$; we consider other distributions in \secref{sec:LWE block fixing sources}.  Our construction uses the code-offset construction~\cite{JW99},\cite[Section 5]{DBLP:journals/siamcomp/DodisORS08} instantiated with a random linear code over a finite field $\Fq$.   Let $\decode_t$ be an algorithm that decodes a random linear code with at most $t$ errors (we will present such an algorithm later, in \label{sec:time main construction}).

@@

The main advantage to analyzing the construction as a whole~(instead of \emph{sketch-then-cextract}) is that security need not depend on the decoding radius $t$.  In \emph{sketch-then-cextract} paradigm, the computational extractor must be able to handle sources with $\Hoo(W| \sketch(W))$.  As shown in \secref{sec:defCompFuzzyExtractors}, this value depends on bounds from error-correcting codes.  In our construction, the decoding radius does not affect the security of the construction.  The decoding radius only impacts the running time of the decoding algorithm.  Conceivably, a computational fuzzy extractor could be built where only correctness depends on $t$~(not efficiency or security).
%However, the construction of a computational fuzzy extractor by 
%The advantage of \consref{cons:informal construction} is that losses due to error correction $O(\log |B_t(\cdot)|)$ and randomness extraction $O(\log 1/\epsilon)$ are not necessary.  In the case where $W$ is small, these losses may leave too few bits for a secure pseudorandom generator.  The work of~\cite{dachman2012computational} shows that these losses in the extract-then-expand construction are in some sense necessary~(the samplable RT conjecture in~\cite{dachman2012computational} has been resolved positively by~\cite{dodis2013key}).

@@

\section{\consref{cons:informal construction} with Nonuniform Sources}
\lnote{what's up with this section/subsection thing?}
\subsection{\consref{cons:informal construction} with Symbol Fixing Sources}
\label{sec:LWE block fixing sources}
In this section, we show the security of~\consref{cons:informal construction} is preserved for certain high entropy distributions.  Showing security of a computational fuzzy extractor for arbitrary high min-entropy distributions is an open problem.  First we recall the notion of a symbol fixing source~(from~\cite{KZ07}): 
\begin{definition}
Let $W = (W_1,..., W_{m+\alpha})$ be a distribution where $W_i$ takes values over $\{0,1\}^b$, we say that a distribution is a $m$-\emph{symbol fixing source} if $H_\infty(W)\geq mb$ and for all $W_i$ one of the following conditions is met:
\begin{itemize}
\item $W_i \overset{d}= U_b$
\item There exists a value $c_i$ such that $\Pr[W_i = c_i] =1$.
\end{itemize}
\end{definition}
%A block fixing source is an extension of a entropy source where each bit is independently random or fixed to larger alphabets.

@@

In the information theoretic setting, Dodis et. al.~\cite{DBLP:journals/siamcomp/DodisORS08} construct fuzzy extractors by using a secure sketch to perform (secure)~error correction and a randomness extractor to convert the distribution to close to uniform.  A similar paradigm seems desirable in the computational case: can a computational fuzzy extractor can be produced by a ``computational'' secure sketch and a randomness extractor~(which convert conditional HILL entropy to pseudorandom bits~\cite[Lemma 5]{DBLP:conf/eurocrypt/HsiaoLR07}).
}
\bibliographystyle{alpha}
\bibliography{crypto}
\appendix
\section{Random Linear Codes Have High Distance}
%\subsection{Notions of Computational Entropy}
%\label{sec:notionsEntropy}
%In this sections we provide some further definitions of computational entropy.  HILL entropy is defined in \secref{sec:defCompFuzzyExtractors}.  HILL entropy generalizes the notion of indistinguishability.  Yao entropy~\cite{DBLP:conf/focs/Yao82a, barak-computational} generalizes the notion of compressibility.  Here we present the computational definition due to~\cite{DBLP:conf/eurocrypt/HsiaoLR07}:
%\begin{definition}
%\label{def:yao entropy}
%For a distribution $X, Z$ we say that $X$ has Yao entropy at least $k$ conditioned on $Z$, denoted $H^{\yao}_{\epsilon, s}(X|Z)\geq k$, if for every pair of circuits $c,d$ of total size at most $s$ with the outputs of $c$ having length $\ell$,
%\[
%\Pr_{(x,z)\leftarrow (X,Z)}[d(c(x,z),z) = x]\leq 2^{\ell-k}+\epsilon
%\]
%\end{definition}
%It should be clear that HILL entropy implies Yao entropy in both the unconditional and conditional case.  In the conditional case it is possible for Yao entropy to be significantly larger than HILL entropy~\cite{DBLP:conf/eurocrypt/HsiaoLR07}.  We also use a less common notion called unpredictability entropy.  This measures an adversaries ability to recover a secret given some side information, see \defref{def:unp entropy}.
%Unpredictability entropy is a notion between HILL and Yao entropy:
%\begin{lemma}~\cite[Lemmas 8, 9]{DBLP:conf/eurocrypt/HsiaoLR07}\label{lem:hillimplyunpimplyyao}
%\[ H^{\hill}_{\epsilon, s}(X|Z)\geq k \Rightarrow H^{\unp}_{\epsilon, s}(X|Z)\geq k\Rightarrow H^{\yao}_{\epsilon, s}(X|Z)\geq k.
%\]
%\end{lemma} 
%\subsection{Unique Witness Relations}

>>>>

\paragraph {Our Negative Results}
We first study (in  \secref{sec:impossCompSecSketch}) whether it could be fruitful to relax the definition of the main building block of a fuzzy extractor, called \emph{secure sketch}.  A secure sketch provides information reconciliation: it  produces a public value $s$ that allows recovery of $w$ from any close value $w'$.  The traditional secrecy requirement of a secure sketch is that $w$ has high min-entropy conditioned on $s$.  This allows the fuzzy extractor of~\cite{DBLP:journals/siamcomp/DodisORS08}  to form the key $r$ by applying a randomness extractor~\cite{nisan1993randomness} to $w$, because randomness extractors get random strings from strings with conditional min-entropy. We call this the ``sketch-and-extract'' construction.

@@

By considering this computational secrecy requirement, we construct the first \emph{lossless} computational fuzzy extractors (\consref{cons:informal construction}), where the derived key $r$ is as long as the entropy of the source $w$. It is uses the code-offset construction~\cite{JW99},\cite[Section 5]{DBLP:journals/siamcomp/DodisORS08} used in prior work, but with two crucial differences.  First, the key $r$ is not extracted from $w$ like in the sketch-and-extract approach; rather $w$ ``encrypts'' $r$ in a way that is decryptable with the knowledge of some close $w'$. Second, the code used is a random linear code, which allows us to use the Learning with Errors~(LWE) assumption due to Regev~\cite{regev2005LWE, regevLWEsurvey} to get a long $r$ out.

Specifically,  we use the  recent result of D\"{o}ttling and M\"{u}ller-Quade~\cite{dottling2012}, which shows the hardness of decoding random linear codes when the error vector comes from the uniform distribution, with each coordinate ranging over a small interval. This allows us to use $w$ as the error vector, assuming it is uniform.  We also use a result of Akavia, Goldwasser, and Vaikuntanathan~\cite{akavia2009},  which says that LWE has many hardcore bits, to hide $r$.

Because we use a random linear code,  our decoding is limited to reconciling a logarithmic number of differences.  Unfortunately, we  cannot utilize the results that improve the decoding radius through the use of trapdoors (such as \cite{regev2005LWE}), because in a fuzzy extractor, there is no secret storage place for the trapdoor. If improved decoding algorithms are obtained for random linear codes, they will improve error-tolerance of our construction.

%We show that this relaxation of the security requirement is  useful.  Specifically, we also define the notion of computational fuzzy extractors, in which the extracted key is required to be pseudorandom---i.e., to look random to any computationally bounded observer.
%We show that our relaxed notion of secure sketches is sufficient to construct computational  fuzzy extractors via a simple and efficient construction, utilizing reconstructive extractors~\cite{barak-computational}.  The resulting fuzzy extractor produces a key that is as long as the entropy of $w$, minus only the loss necessary in any extractor construction.
%\lnote{it would be good to find another application}

In \secref{sec:LWE block fixing sources}, we are able to relax the assumption that $w$ comes from the uniform distribution, and instead allow $w$ to come from a symbol-fixing source~\cite{KZ07} (each dimension is either uniform or fixed). This relaxation follows from our results about the hardness of LWE when samples have a fixed~(and adversarially known) error vector, which may be of independent interest~(\thref{thm:blockLWE}).

\paragraph{An Alternative Approach}  Computational extractors~\cite{krawczyk2010cryptographic, barak2011leftover, dachman2012computational} have the same goal of obtaining a pseudorandom key $r$ from a source $w$ in the setting without errors.  They can be constructed, for example, by applying to pseudorandom generator to the output of an information-theoretic extractor.  One way to build a computational \emph{fuzzy} extractor is by using a computational extractor instead  of the information-theoretic extractor in the sketch-and-extract construction of  \cite{DBLP:journals/siamcomp/DodisORS08}.   However, this approach is possible only if conditional min-entropy of $w$ conditioned on the sketch $s$ is high enough.  
%We ask whether it is possible to avoid this loss by making whole process ``computational.''
We compare the two approaches in \secref{sec:prg based comparison}.

@@

%In \secref{sec:impossCompSecSketch} we will use a slightly weaker notion, in which \rec is allowed to fail with small probability.\lnote{Can we add a comment that this corresponds to errors that can depend on the input but not the sketch? It's not obvious from the way this is written.} \lnote{can we just have a single definition --- the one below --- instead of two?}
%\begin{definition}
%We say that pair of procedures $(\sketch, \rec)$ is $(\mathcal{M}, m, \tilde{m}, t, \epsilon)$ is an \emph{approximately correct secure sketch} if \defref{def:secure sketch} holds with correctness replaced with the following property, $\forall w, w'$ such that $\dis(w, w')\leq t$:
%\[
%\Pr[\rec(w', \sketch(w)) = w]\geq 1-\epsilon.
%\] where the probability is taken over the coins of \sketch and \rec.
%\end{definition}
\begin{lemma}%\protect{\cite[Lemma 4.1]{DBLP:journals/siamcomp/DodisORS08}}
\label{lem:fuzzy ext construction}
Assume $(\sketch, \rec)$ is an $(\mathcal{M}, m, \tilde{m}, t)$-secure sketch with error $\delta$, and let $\ext:\mathcal{M}\times \zo^d \rightarrow \zo^\ell$ be a $(\tilde{m}, \epsilon)$-average-case extractor.  Then the following $(\gen, \rep)$ is an $(\mathcal{M}, m, \ell, t, \epsilon)$-fuzzy extractor with error $\delta$:
\begin{itemize}
\item $\gen(w):$ set $P=(\sketch(w), x), R=\ext(w;x)$, and output $(R,P)$.
\item $\rep(w', (s, x)):$ recover $w=\rec(w',s)$ and output $R=\ext(w;x)$.
\end{itemize}
\end{lemma}
The main parameter we will be concerned with is the entropy loss of the construction.  In this paper, we ask whether a smaller entropy loss can be achieved by considering a fuzzy extractor with a computational security requirement.  We relax the security requirement of \defref{def:fuzzy extractor} to require a pseudorandom output instead of a truly random output.

@@

%=============================================================
\section{Impossibility of Computational Secure Sketches}
\label{sec:impossCompSecSketch}
In this section, we consider whether it is possible in build a secure sketch that retains significantly more computational than information-theoretic entropy.  We consider two different notions for computational entropy, and for both of them show that corresponding secure sketches are subject to the same upper bounds as those for information-theoretic secure sketches. Thus, it seems that relaxing security of sketches from information-theoretic to computational does not help.

@@

\begin{theorem}
\label{thm:imp of unp entropy}
Let $Z$ be an alphabet $(\sketch, \rec)$ be an unpredictability-entropy $(Z^n, m, \tilde{m}, t)$-secure sketch that is $(\epsilon, s_{sec})$-secure with error $\delta$.  If $s_{sec} \geq t(|\rec|+n\log |Z|)$, then $\tilde{m}\leq n\log |Z| - \log |B_t(\cdot)| + \log(1-\epsilon -t\delta)$.
\end{theorem}
In particular, if the input is uniform, the entropy loss is about $\log |B_t(\cdot)|$.  As mentioned at the beginning of~\secref{sec:impossCompSecSketch}, essentially the same entropy loss can be achieved with information-theoretic secure sketches, by using the randomized code-offset construction. However, it is conceivable that unpredictability entropy secure sketches could achieve lower entropy loss with greater efficiency for some parameter settings.

@@

In this section we describe our main construction.  In this section, we consider only a uniform source $W$; we consider other distributions in \secref{sec:LWE block fixing sources}.  Our construction uses the code-offset construction~\cite{JW99},\cite[Section 5]{DBLP:journals/siamcomp/DodisORS08} instantiated with a random linear code over a finite field $\Fq$.   Let $\decode_t$ be an algorithm that decodes a random linear code with at most $t$ errors (we will present such an algorithm later, in \secref{sec:time main construction}).

@@

The main advantage to analyzing the construction as a whole~(instead of \emph{sketch-then-cextract}) is that security need not depend on the decoding radius $t$.  In \emph{sketch-then-cextract} paradigm, the computational extractor must be able to handle sources with $\Hoo(W| \sketch(W))$.  As shown in \secref{sec:impossCompSecSketch}, this value depends on bounds from error-correcting codes.  In our construction, the decoding radius does not affect the security of the construction.  The decoding radius only impacts the running time of the decoding algorithm.  Conceivably, a computational fuzzy extractor could be built where only correctness depends on $t$~(not efficiency or security).
%However, the construction of a computational fuzzy extractor by 
%The advantage of \consref{cons:informal construction} is that losses due to error correction $O(\log |B_t(\cdot)|)$ and randomness extraction $O(\log 1/\epsilon)$ are not necessary.  In the case where $W$ is small, these losses may leave too few bits for a secure pseudorandom generator.  The work of~\cite{dachman2012computational} shows that these losses in the extract-then-expand construction are in some sense necessary~(the samplable RT conjecture in~\cite{dachman2012computational} has been resolved positively by~\cite{dodis2013key}).

@@

\section{Computational Fuzzy Exactor with Nonuniform Sources}
\label{sec:LWE block fixing sources}
In this section, we show the security of~\consref{cons:informal construction} is preserved for certain high entropy distributions.  Showing security of a computational fuzzy extractor for arbitrary high min-entropy distributions is an open problem.  First we recall the notion of a symbol fixing source~(from~\cite{KZ07}): 
\begin{definition}
Let $W = (W_1,..., W_{m+\alpha})$ be a distribution where $W_i$ takes values over $\{0,1\}^b$, we say that a distribution is a $m$-\emph{symbol fixing source} if $H_\infty(W)\geq mb$ and for all $W_i$ one of the following conditions is met:
\begin{itemize}
\item $W_i \overset{d}= U_b$
\item There exists a value $c_i$ such that $\Pr[W_i = c_i] =1$.
\end{itemize}
\end{definition}
%A block fixing source is an extension of a entropy source where each bit is independently random or fixed to larger alphabets.

@@

In the information theoretic setting, Dodis et. al.~\cite{DBLP:journals/siamcomp/DodisORS08} construct fuzzy extractors by using a secure sketch to perform (secure)~error correction and a randomness extractor to convert the distribution to close to uniform.  A similar paradigm seems desirable in the computational case: can a computational fuzzy extractor can be produced by a ``computational'' secure sketch and a randomness extractor~(which convert conditional HILL entropy to pseudorandom bits~\cite[Lemma 5]{DBLP:conf/eurocrypt/HsiaoLR07}).
}
\bibliographystyle{alpha}
\bibliography{crypto}
\appendix
\section{Random Linear Codes Have High Distance}
%\subsection{Notions of Computational Entropy}
%\label{sec:notionsEntropy}
%In this sections we provide some further definitions of computational entropy.  HILL entropy is defined in \secref{sec:impossCompSecSketch}.  HILL entropy generalizes the notion of indistinguishability.  Yao entropy~\cite{DBLP:conf/focs/Yao82a, barak-computational} generalizes the notion of compressibility.  Here we present the computational definition due to~\cite{DBLP:conf/eurocrypt/HsiaoLR07}:
%\begin{definition}
%\label{def:yao entropy}
%For a distribution $X, Z$ we say that $X$ has Yao entropy at least $k$ conditioned on $Z$, denoted $H^{\yao}_{\epsilon, s}(X|Z)\geq k$, if for every pair of circuits $c,d$ of total size at most $s$ with the outputs of $c$ having length $\ell$,
%\[
%\Pr_{(x,z)\leftarrow (X,Z)}[d(c(x,z),z) = x]\leq 2^{\ell-k}+\epsilon
%\]
%\end{definition}
%It should be clear that HILL entropy implies Yao entropy in both the unconditional and conditional case.  In the conditional case it is possible for Yao entropy to be significantly larger than HILL entropy~\cite{DBLP:conf/eurocrypt/HsiaoLR07}.  We also use a less common notion called unpredictability entropy.  This measures an adversaries ability to recover a secret given some side information, see \defref{def:unp entropy}.
%Unpredictability entropy is a notion between HILL and Yao entropy:
%\begin{lemma}~\cite[Lemmas 8, 9]{DBLP:conf/eurocrypt/HsiaoLR07}\label{lem:hillimplyunpimplyyao}
%\[ H^{\hill}_{\epsilon, s}(X|Z)\geq k \Rightarrow H^{\unp}_{\epsilon, s}(X|Z)\geq k\Rightarrow H^{\yao}_{\epsilon, s}(X|Z)\geq k.
%\]
%\end{lemma} 
%\subsection{Unique Witness Relations}
